---
title: "Introduction to ECLIPSE"
subtitle: "ECLIPSE: Enhanced CLonotypic Inference via Prediction of 
Single-cell Expression"
author:
- name: Ethan C. Burns
  email: ethan.burns@yale.edu
  affiliation: Yale School of Medicine, New Haven, CT, USA
- name: David A. Braun
  affiliation: Yale School of Medicine, New Haven, CT, USA
package: ECLIPSE
abstract: |
  This vignette displays how to use the `ECLIPSE` package for the analysis of 
  single-cell T cell receptor sequencing data (scTCR-seq). The package utilizes
  Seurat objects and `cell ranger vdj/multi` output files as its main inputs. 
  Core functionalities include the annotation of T cell clones, the
  modification of `cell Ranger vdj/multi` output files to reflect clones with 
  extra TCR chains, and the filtering of cells with excessive TCR chains from 
  Seurat objects. Please refer to our 
  [pre-print](https://www.biorxiv.org/content/10.64898/2026.02.18.706444v1)
  for more details.
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{ECLIPSE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    dev = "ragg_png"
)
```

# Introduction
Traditional analysis of single-cell TCR sequencing (scTCR-seq) data defines 
T cell clones based on the paired CDR3 sequences of the 
TCR$\alpha$ and TCR$\beta$ chains. However, as >50% of cells in some samples 
possess only 1 or >2 TCR chains, this approach is ill-suited to interpret the 
clonotypes of a substantial proportion of cells.

ECLIPSE was developed to address this limitation of standard scTCR-seq 
analysis. By analyzing chain pairings across a sample, it is able to resolve 
clonal ambiguity, allowing for the prediction of chains lost of dropout and the 
deciphering of true biological expression of extra TCR chains from technical 
artifacts. With less missing data and larger clone sizes, users are then able 
to make best use of their scTCR-seq data. The ECLIPSE workflow is also quite 
streamlined, typically utilizing just 2 core functions (`ECLIPSE` --> 
`tcrDoubletDetect`). After completing these two functions, users are ready
for convenient downstream analysis with the `r Biocpkg("scRepertoire")` 
package.

This vignette will walk through the basic workflow of the ECLIPSE package.
Please refer to our 
[pre-print](https://www.biorxiv.org/content/10.64898/2026.02.18.706444v1)
for more details about this method.

# Installation and loading

We need to first load three packages to run the vignette:
```{r loadLibraries, results = "hide", message = FALSE, warning = FALSE}
library(remotes) # only needed to insall the package
library(tidyverse) # needed for data wrangling (or use base R instead)
library(scRepertoire) # needed for downstream analysis
```

Please use the following code to download `ECLIPSE` from GitHub and load it:
```{r githubInstall, eval = FALSE}
install_github("eburns10/ECLIPSE")
```

```{r loadECLIPSE}
library(ECLIPSE)
```

```{r loadKnitr, echo = FALSE, message = FALSE}
require(knitr)
```


# Test dataset
ECLIPSE includes a small test dataset of 1,000 cells from 2 patients (S3 and
S15) with clear cell renal cell carcinoma (derived from [Braun et al., 2021](
https://doi.org/10.1016/j.ccell.2021.02.013)). Each patient has two samples: 
one from tumor tissue and one from matched adjacent normal tissue. To load the
dataset in, we can use the following code:
```{r loadSeuratObject}
data("vignette_object")
```


# Preparing inputs for `ECLIPSE`
2 main inputs are required for `ECLIPSE`: a Seurat object that contains 
scRNA-seq data and contig files from `cell ranger vdj/multi` (preferably 
`all_contig_annotations.csv` files). 

## Contig files
The first input is contig files from `cellranger multi` or `cellranger vdj`, 
specifically `all_contig_annotations.csv` or `filtered_contig_annotations.csv`
files. **We strongly recommend using all_contig_annotations** files as the 
filtered contig files remove some biologically true contigs (ex.
TCR$\alpha$ chains that use the TRDV1 gene segment, [see here](
https://doi.org/10.3389/fimmu.2023.1286688)) as well as non-productive 
contigs. While these non-productive chains may not influence T cell antigen 
specificity, their presence in the contig files allows for enhanced fidelity  
of clonal tracking by `ECLIPSE`.

To feed the contig files into `ECLIPSE`, there are two options:

### Option 1: use folder paths
For this option, a vector of folder paths that contain contig files is needed. 
`ECLIPSE` will find the contig files within the folders you 
provide, however it requires that each folder only contains 1 set of contig
files (i.e. you cannot put the contig files from each sample into 1 folder).
In order to create a vector of folder paths, we can use the `list.dirs()`
function:

```{r getFolderPath}
# Finding where the vignette data for ECLIPSE is stored
vignette_dir <- system.file("extdata/vignette", package = "ECLIPSE")

# Generates the vector of folder paths and excludes the current directory
tcr_folder_paths <- list.dirs(path = vignette_dir, full.names = TRUE)[-1]
print(tcr_folder_paths)
```


### Option 2: use contig file paths
The second option also requires a vector of paths, but instead you specify
the full path to the contig file. This option would be better if you have many
contig files from different samples in one folder. 
To make the vector, here we can use `list.files()`:

```{r getFilePath}
# Generates the vector of file paths
contig_paths <- list.files(path = vignette_dir,
                           pattern = "all_contig", 
                           recursive = TRUE, 
                           full.names = TRUE)
contig_paths
```


## Seurat object meta.data
4 arguments of `ECLIPSE` must be provided column names from the Seurat 
meta.data: `original_barcode`, `batch`, `donor`, and `group`. We will create 
these columns now.

### Column to pass to `original_barcode`
This column must contain the original 10X barcode for each cell. Because many
scRNA-seq analysis pipelines add prefixes/suffixes to cell barcodes, this 
column is required so that the TCR contigs (which contain the original 10X
barcodes) can be matched to their respective cell in the Seurat object. 
Typically, barcodes will end with -1, but this may not always be true, such 
as if `cellranger aggr` was used. Always check your raw contig and counts 
files to be sure! 

For our dataset, the raw contig files have cell barcode suffixes of -1, so 
let's look at the barcodes in our Seurat object to see if they will match:
```{r checkBarcodes}
# Existing cell barcodes. They aren't the original!
head(rownames(vignette_object[[]]))
```

As you can see, they do not match. Let's change the suffixes to -1 and make a 
new column called `cell_barcode` that we can pass to the `original_barcode` 
argument of ECLIPSE:
```{r makeBarcodeColumn}
# Making a column with these barcodes where the suffix is -1
vignette_object$cell_barcode <- str_replace(rownames(vignette_object[[]]),
                                            pattern = "-.*$", 
                                            replacement = "-1")
head(vignette_object$cell_barcode) %>% as.vector()
```

### Column to pass to `batch`
This column tells `ECLIPSE` which contig file holds the TCR information for
each cell in the Seurat object. The position of the contig file in the vector
must be provided here (ex. 2, but not "2").
```{r checkContigPaths}
# Looking at the vector of file_paths
contig_paths
```

For example, `contig_paths[1]` holds a path for the TCR contigs for a cells
from the normal tissue of S15, so any cell from that sample must have 1 in 
this new column. We can make this column, here called `index`, as such:
```{r makeBatchColumn}
# Since normal tissue from sample 15 is first (S15N), it is 1.
# Tumor tissue from 15 is next (S15T), so it is 2, etc...
vignette_object[[]] <- vignette_object[[]] %>% 
                           mutate(index = case_when(
                               Sample == "S15_N" ~ 1,
                               Sample == "S15_T" ~ 2,
                               Sample == "S3_N" ~ 3,
                               Sample == "S3_T" ~ 4))
    
# Checking the column
table(vignette_object$index)
```

### Column to pass to `donor`
Next, we need to tell `ECLIPSE` which donor (or mouse) each cell was from. As 
`ECLIPSE` calls clonotypes by utilizing chain pairing across a repertoire, the 
more cells that are available, the more accurate this prediction will be 
(regardless of time/condition/treatment). For this reason, a column is 
required to provide the donor/mouse. Even if we want to compare S3 tumor vs. 
normal and S15 tumor vs. normal downstream, here we must provide S3 and S15. 
Fortunately, this column is already created for us, and it should be already 
present in most real-world uses. Let's look at this column, called `donor`:
```{r checkDonorColumn}
# Show donor column
table(vignette_object$donor)
```

### Column to pass to `group`
Lastly, we need to tell `ECLIPSE` how we want the final clonal statistics 
(i.e. clone cell count, frequency, etc..) to be calculated after clones are 
called. For our example, either S15/S3 or S3T/S3N/S15T/S15N would be equally 
valid. If a clone is only 1 cell, then if we provide the donor, the 
frequency will be 1/500 vs. if we provide the sample, then it will be
1/250. **Please note that this column can and often will be the same as the **
**column passed to donor**. Let's choose by sample here, which again is 
already created for us through the column `Sample` (and typically will be 
created in most real cases):
```{r checkSampleColumn}
# Show Sample column
table(vignette_object$Sample)
```

# Running the `ECLIPSE` function

## Description of function
Now we are ready to run the `ECLIPSE` function to analyze our scTCR-seq data. 
Briefly, this function does the following:

1. Reading in contig data
2. Filtering contigs (i.e. must be in Seurat, high_confidence, not duplicated,
TCR$\alpha$ or TCR$\beta$, have a valid CDR3, etc...)
3. Calling `findThreeChainClones` to find clones with biological expression of 
3 total TCR chains and modifying the contigs to address this
4. Utilizing the implementation of the EM algorithm in 
`r Biocpkg("VDJdive")`'s clonoStats() function. This function analyzes chain 
pairings across a sample to determine the probability of each potential chain 
pairing for cells possessing only 1 or >2 total TCR chains 
4. Determining the final clonotype (stored in a column called `CTaa`). If 
it is highly likely the cell is in 1 clone, then that clone is used. If no 
clonotype is predicted with high confidence, then clonal prediction is not 
performed, and the final clonotype is instead determined by 
`scRepertoire::combineTCR()` using parameters passed to `ECLIPSE` from the 
`combineTcrArgs` argument (see supplemental information section)
5. Formatting the object for downstream analysis with 
`r Biocpkg("scRepertoire")`

For a full list of columns added by ECLIPSE and their description, please 
refer to the supplemental information section.

## Running `ECLIPSE`
To run `ECLIPSE`, there are two options, depending on whether we generated a 
vector of folders containing contigs or contig file paths. If we want to use
a vector of folder paths, we can use the following code:
```{r runEclipseFolder}
# Option 1: Running with paths of folders containing contigs
object_after_ECLIPSE <- ECLIPSE(contig_type = "all",
                                contig_folder_paths = tcr_folder_paths,
                                seurat_object = vignette_object,
                                batch = "index",
                                donor = "donor",
                                group = "Sample",
                                original_barcode = "cell_barcode",
                                verbose = FALSE)
```

Note that we use the `contig_folder_paths` argument, omit the 
`contig_file_paths` argument, and `contig_type = "all"`. Together, this tells 
`ECLIPSE` to search for `all_contig_annotations.csv` files in the folders that 
we provide.

If we want to provide TCR contig file paths, we can use the 
`contig_file_paths` argument, omit the `contig_folder_paths` argument, and 
`contig_type = "manual"`:
```{r runEclipseFile, fig.alt = "Plot from ECLIPSE"}
# Option 2: Running with contig paths
object_after_ECLIPSE <- ECLIPSE(contig_type = "manual",
                                contig_file_paths = contig_paths,
                                seurat_object = vignette_object,
                                batch = "index",
                                donor = "donor",
                                group = "Sample",
                                original_barcode = "cell_barcode")
```

As seen above, by default (verbose = TRUE), `ECLIPSE` will print some key 
stats during filtering as well as a QC plot that shows how well clonal 
prediction went. For this experiment, most cells (shown in orange) received 
their final clonotype by prediction. Many of these cells also were fully 
assigned to one clone by the EM algorithm (i.e. y = 1), so overall the 
analysis went quite well. 

Let's take a look at the final clonotypes that `ECLIPSE` decided on, now 
stored in the `CTaa` column of the Seurat meta.data:
```{r checkClonotypes}
# Final ECLIPSE clonotype
head(object_after_ECLIPSE$CTaa) %>% as.vector()
```

We can also compare these final clonotypes to what traditional analysis 
would have determined, now stored in the `scR_CTaa` column: 
```{r checkTraditionalClonotypes}
# Clonotype from traditional analysis
head(object_after_ECLIPSE$scR_CTaa) %>% as.vector()
```

ECLIPSE also provides clone sizes and clonal proportions in the 
`clonalFrequency` and `clonalProportion columns,` respectively, with matched 
columns having a scR_prefix to show what would happen with traditional 
analysis.

# Running `tcrDoubletDetect`
Now that the `ECLIPSE` function has been ran, we should now run the 2nd core 
ECLIPSE function, `tcrDoubletDetect`, to filter out cells with excessive TCR 
chains (these cells are likely a mix of doublets and cells with TCR 
contamination). Here we remove any cells with >2 TCR$\alpha$ chains, >2 
TCR$\beta$ chains, or >4 total ($\alpha$ + $\beta$) chains:
```{r runDoubletDetect, fig.alt = "Plots from tcrDoubletDetect"}
object_after_ECLIPSE <- tcrDoubletDetect(object_after_ECLIPSE,
                                         singleChainLimit = 2,
                                         totalChainLimit = 4)
```
This function by default also prints plots that may be useful in setting
additional filtering thresholds. Here we can see that cells with 4 total TCR 
chains have much higher RNA counts, so we may want to remove these cells or 
any cells with ~ >7k nCount_RNA as they may be enriched for doublets. 

# Downstream analysis with `r Biocpkg("scRepertoire")`
At this point we're ready for downstream analysis. ECLIPSE generates Seurat
objects formatted for use with the popular `r Biocpkg("scRepertoire")` 
package, so functions introduced in that package typically can be used here 
as well. One key thing to note is that since `ECLIPSE` predicts clones based 
on CDR3 amino acid sequences, we must always use the argument 
`cloneCall = "aa"` in scRepertoire functions. 

Shown below are a couple potential types of downstream analysis made possible
with `r Biocpkg("scRepertoire")`. For example, we can look at clonal sharing 
across samples:
```{r runCompareClonotypes, fig.alt = "Plot showing clone sharing"}
clonalCompare(object_after_ECLIPSE,
              top.clones = 10,
              graph = "alluvial",
              cloneCall = "aa",
              group.by = "Sample") +
    theme(legend.position = "none")

```

We can also visualize which samples have a larger fraction of expanded 
clonotypes:
```{r runClonalHomeostasis, fig.alt = "Plot showing clone sizes"}
clonalHomeostasis(object_after_ECLIPSE,
                  cloneCall = "aa",
                  group.by = "Sample")
```

Or we can look at repertoire diversity across samples:
```{r runClonalDiversity, fig.alt = "Plot showing TCR diversity"}
clonalDiversity(object_after_ECLIPSE,
                metric = "shannon",
                cloneCall = "aa",
                group.by = "Sample")
```

# Conclusions
We hope this vignette is a helpful introduction to using ECLIPSE. If you have 
any questions, please email ethan.burns [@] yale.edu or use our issues 
[page](https://github.com/eburns10/ECLIPSE/issues). For more details about 
our method, please refer to our 
[pre-print](https://www.biorxiv.org/content/10.64898/2026.02.18.706444v1).


# Supplemental information

## Helper function: `findThreeChainClones`
While not discussed above, ECLIPSE contains a third function that may
may be of use. This function was designed to help the `ECLIPSE` function
find clones with biological expression of 3 TCR chains, but can be also
used to see which clones were called with 3 TCR chains and how that occurred.
To see these clones we can use the code below. First, we generate QC filtered
contigs using the `returnFilteredContigs` argument of `ECLIPSE`:
```{r makeContigs}
postqc_contigs <- ECLIPSE(contig_type = "all",
                          contig_folder_paths = tcr_folder_paths,
                          seurat_object = vignette_object,
                          batch = "index",
                          donor = "donor",
                          group = "Sample",
                          original_barcode = "cell_barcode",
                          verbose = FALSE,
                          returnFilteredContigs = TRUE)
```

Next, we filter on a sample we are interested in, and pass the contigs onto
`findThreeChainClones` using ` mode = "show"` and `chain = "a"`
to show clones with 2 TCR$\alpha$ chains:
```{r runFindThreeChainClones}
s15_contigs <- postqc_contigs %>% filter(donor == "S15")

three_chain_clones <- findThreeChainClones(s15_contigs,
                                           chain = "a",
                                           mode = "show",
                                           printStats = TRUE)
```

This provides a list containing:

1. A table of all clones with 2 TCR$\alpha$ chains after filtering
```{r getOutput1}
three_chain_clones[[1]]
```

2. Pre-filtering table that shows how some of these clones were called
```{r getOutput2}
three_chain_clones[[2]]
```




## `ECLIPSE` arguments
Here are additional arguments for the `ECLIPSE` function:

1. `verbose`: by default `TRUE`. Prints stats such as QC filtering efficacy, 
how many clones were called with 3 chains, and final percentages of cells 
with a TCR called
2. `returnFilteredContigs`: by default `FALSE`. `TRUE` would return the 
contigs after QC filtering
3. `allowDoubleChains`: by default `"both"`, whereby clones can have 2 
TCR$\alpha$ or TCR$\beta$ chains. If `"alpha"` or `"beta"` is provided, only 
clones with 2 of that type will be permitted. If `"neither"` is provided, 
then only clones with 2 chains will be called
4. `combineTcrArgs`: a list that determines how `scRepertoire::combineTCR()` 
is called. As `ECLIPSE` runs `combineTCR/combineExpression` to allow users to
combine the predicted `ECLIPSE` clonotype with a traditionally-defined 
clonotype, this 
tells which arguments `combineTCR` should be ran with. Specifically, it tells 
the 4 arguments 1) `removeNA`, 2) `removeMulti`, 3) `filterMulti`, 4) 
`filterNonproductive` in that order. If `ECLIPSE` cannot call a clonotype 
with high accuracy, the clonal annotation from this will be used as the final 
clonotype. By default, this is `list(FALSE, FALSE, FALSE, TRUE)` which calls 
`combineTCR` with its default arguments (i.e. no non-productive contigs are 
retained but clones with 1 TCR chain or >2 are allowed)
5. `detailedEclipseStats`: by default `FALSE`. This adds information about the
EM algorithm assignment to the Seurat object (i.e. how likely is the cell
in the clone that is listed)
6. `format`: by default `"Blank"`. This tells how missing data is stored in 
contig files. Typically this is `""` so `"Blank"` should be specified, 
however older versions of Cell Ranger use `"None"` instead. If this is the 
case, please provide `"None"` here.

```{r makeArgsTable, echo = FALSE}
args <- c("`verbose`",
            "`returnFilteredContigs`",
            "`allowDoubleChains`",
            "`combineTcrArgs`",
            "`detailedEclipseStats`",
            "`format`")

descriptions <- c(
    paste(
        "By default `TRUE`. Prints stats such as QC filtering efficacy,",
        "how many clones were called with 3 chains, and final percentages",
        "of cells with a TCR called"
    ),
    paste(
        "By default `FALSE`. `TRUE` would return the contigs after",
        "QC filtering"
    ),
    paste(
        "By default `\"both\"`, whereby clones can have 2 TCR$\\alpha$ or",
        "TCR$\\beta$ chains. If `\"alpha\"` or `\"beta\"` is provided, only",
        "clones with 2 of that type will be permitted. If `\"neither\"` is", 
        "provided, then only clones with 2 chains will be called"
    ),
    paste(
        "A list that determines how `scRepertoire::combineTCR()` is called.",
        "As `ECLIPSE` runs `combineTCR/combineExpression` to allow users to",
        "combine the predicted `ECLIPSE` clonotype with a",
        "traditionally-defined clonotype, this tells which arguments",
        "`combineTCR` should be run with. Specifically, it tells the 4",
        "arguments 1) `removeNA`, 2) `removeMulti`, 3) `filterMulti`, and 4)", 
        "`filterNonproductive` in that order. If `ECLIPSE` cannot call a",
        "clonotype with high accuracy, the clonal annotation from this will",
        "be used as the final clonotype. By default, this is",
        "`list(FALSE, FALSE, FALSE, TRUE)`, which calls `combineTCR` with its",
        "default arguments (i.e. no non-productive contigs are retained but",
        "clones with 1 TCR chain or >2 are allowed)"
    ),
    paste(
        "By default `FALSE`. This adds information about the EM algorithm",
        "assignment to the Seurat object (i.e. what proportion of the cell",
        "was assigned to the clone listed, etc...)"
    ),
    paste(
        "By default `\"Blank\"`. This tells how missing data is stored in", 
        "contig files. Typically this is `\"\"` so `\"Blank\"` should be",
        "specified, however older versions of Cell Ranger use `\"None\"`",
        "instead. If this is the case, please provide `\"None\"` here"
    )
)

argsTable <- data.frame("Argument" = args,
                        "Description" = descriptions)
kable(argsTable, caption = "Arguments for the ECLIPSE function")
```


## Seurat object columns description

1. `highConfChains`: all TCR chains that were sequenced in the cell and passed
QC filtering by ECLIPSE. Helpful to compare to final clone (`CTaa`) used for 
analysis
2. `scR_CTgene`, `scR_CTnt`, `scR_CTaa`, `scR_CTstrict`, 
`scR_clonalProportion`, `scR_clonalFrequency`, `scR_cloneSize`: columns 
added by `scRepertoire::combineExpression`, see `r Biocpkg("scRepertoire")` 
documentation for details. These provide information about the cell if it was 
called with scRepertoire, and they can be useful in comparing to the final 
clone (`CTaa`) used for analysis. Please note the scR_ prefix is added so 
`ECLIPSE` can add other columns with the same name that can interact with 
scRepertoire
3. `cloneCallSource`: whether the final clonotype listed (in `CTaa`) is from 
VDJidve's prediction or from traditional analysis from scRepertoire
4. `CTaa`, `clonalProportion`, `clonalFrequency`, `cloneSize`: columns in the 
style of what would be added by `scRepertoire::combineExpression`. `CTaa` 
stores the final clonotype of the cell
5. `cell_unconventional_subset`, `clone_unconventional_subset`: 
`cell_unconventional_subset` tells if the chains detected suggest the cell 
is a iNKT cell or MAIT cell. `clone_unconventaional_subset` gives similar 
information, but uses the final clonotype of the cell instead of what chains 
were detected in that particular cell (and therefore will capture more cells 
as being a iNKT or MAIT cell). For both of these columns, `ECLIPSE` searches 
for 3 subsets: 1) iNKT cells (must have TRAV10 + TRAJ18 + TRBV25-1), 2) 
conventional MAIT cells (must have TRAV1-2 + TRAJ33)), and 3) unconventional 
MAIT cells (must have TRAV1-2 + TRAJ12 or TRAV1-2 + TRAJ20). Please note that 
these are for human samples only, murine T cells use different gene segments 
that are not searched for by ECLIPSE.

```{r makeColsTable, echo = FALSE}
cols <- c("`highConfChains`",
            paste("`scR_CTgene`, `scR_CTnt`, `scR_CTaa`, `scR_CTstrict`,",
                "`scR_conalProportion`, `scR_clonalFrequency`,",
                "`scR_cloneSize`"),
            "`cloneCallSource`",
            "`CTaa`, `clonalProportion`, `clonalFrequency`, `cloneSize`",
            "`cell_unconventional_subset`, `clone_unconventional_subset`",
           "`EM_max_prop`, `EM_2ndmax_prop`, `EM_max_2ndmax_ratio`")

descriptions <- c(
    paste("all TCR chains that were sequenced in the cell and passed QC",
        "filtering by ECLIPSE. Helpful to compare to final clone (`CTaa`)",
        "used for analysis"),
    paste("columns added by `scRepertoire::combineExpression`, see",
        BiocStyle::Biocpkg("scRepertoire"),
        "documentation for details. These",
        "provide information about the cell if it was called with",
        BiocStyle::Biocpkg("scRepertoire"),
        "and can be useful in comparing to the final clone",
        "(`CTaa`) used for analysis. Please note the \"scR_ prefix\" is added",
        "so ECLIPSE can add other columns with the same name that can",
        "interact with scRepertoire"),
    paste("whether the final clonotype list (in `CTaa`) is from",
        "the prediction of ",
        BiocStyle::Biocpkg("VDJdive"),
        "or from traditional analysis from",
        BiocStyle::Biocpkg("scRepertoire")),
    paste("columns in the style of what would be added by",
        "`scRepertoire:::combineExpression`. However, `CTaa` here stores",
        "the final clonotype of the cell as predicted by `ECLIPSE`"),
    paste("`cell_unconventional_subset` tells if the chains detected suggest",
        "the cell is a iNKT or MAIT cell. `clone_unconventional_subset` gives",
        "similar information, but uses the final clonotype of the cell",
        "instead of what chains were detected in that particular cell (and",
        "therefore will capture more cells as being a iNKT or MAIT cell). For",
        "both of these columns, `ECLIPSE` searches for 3 subsets: 1) iNKT",
        "cells (must have TRAV10 + TRAJ18 + TRBV25-1), 2) conventional MAIT",
        "cells (must have TRAV1-2 + TRAJ33), and 3) unconventional MAIT cells",
        "(must have TRAV1-2 + TRAJ12 or TRAV1-2 + TRAJ20. Please note that",
        "these are for human samples only, murine T cells use different gene",
        "segments that are not searched for by `ECLIPSE`"),
    paste("the proportion of the cell assigned by the EM algorithm to the",
          "most and 2nd most likely clone. The ratio of these two proportions",
          "(biggest/2nd biggest) is also provided. These columns are only",
          "included if `detailedEclipseStats == TRUE`")
)

colsTable <- data.frame("Column" = cols,
                        "Description" = descriptions)
kable(colsTable, 
      caption = "Columns made by the ECLIPSE function", 
      escape = FALSE)
```

# Session info
```{r checkSessionInfo}
sessionInfo()
```
